---
title: "Group 38 CA pollution burden"
author: "Conner Smith and Samantha Kuglen"
date: "11/12/2021"
output: 
  html_document: 
    toc: yes
    toc_depth: 2
    theme: flatly
---

## **1. Overview**

This report analyzes the differences in pollution burdens based on racial makeup of different areas from 2010 census data. This uses data from the California Office of Environmental Health Hazard Assessment (OEHHA)'s [CalEnviroScreen tool](https://oehha.ca.gov/calenviroscreen).

## **2. Setup**
```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(hexbin)
```

```{r}
# 2A: Read in data 
ces <- read_csv(here("data", "ces3demographics.csv")) %>% 
  clean_names()
```

## **3. Pollution Burden by White Population Proportion** 

```{r}
# 2B: Create a scatterplot graph showing percentage of white population versus pollution score. 
# 2C: Update graph type to hexbin, add appropriate axis labels
# 2D: Changed theme to minimal and added color gradient 
# 2E: Add figure caption, update code chunk options
# 2F: Added summary below graph 

ggplot(data = ces, aes(x = white_percent, y = ces_3_0_score)) +
  geom_hex() +
  scale_fill_gradient(low = "thistle1", high = "slateblue4") +
  theme_minimal() +
  labs(x = "White Population (%)", 
       y = "Pollution Burden Score",
       title = "Cumulative Pollution Burden for White Population in California") +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend("Count")) # update legend title
```

**Figure 1.** *CalEnviroScreen 3.0 aggregate score of pollution burden by proportion of white people in the population. The count refers to the number of census tracts that fall in a given bin on the graph. Darker colors indicate a higher number of tracts per bin. Data: 2010 United States Census and OEHHA.*

This graph shows a clear negative correlation between the proportion of white people in a census tract and the aggregate pollution burden. Census tracts with white population share above 75 percent showed pollution scores in the lowest quartile of the range on average. The declining pollution burden with increasing whiteness in a given census tract provides evidence for the fact that low-income communities face higher exposure to environmental harm.

# **4. References**

California Office of Environmental Health Hazard Assesment. CalEnviroScreen 3.0. (2018). https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30
