---
title: "Group 38 CA pollution burden"
author: "Conner Smith and Samantha Kuglen"
date: "11/12/2021"
output: html_document
---

## **Overview**

This report analyzes the differences in polluton burdens based on racial makeup of different areas. This uses data from the California Office of Environmental Health Hazard Assessment (OEHHA)'s [CalEnviroScreen tool](https://oehha.ca.gov/calenviroscreen).

## Setup
```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(hexbin)
```

```{r}
# 2A: Read in data 
ces <- read_csv(here("data", "ces3demographics.csv")) %>% 
  clean_names()
```

## **2. Pollution Burden by Race** 

```{r}
# 2B: Create a scatterplot graph showing percentage of white population versus pollution score. 
# 2C: Update graph type to hexbin, add appropriate axis labels
# 2D: Changed theme to minimal and added color gradient 
# 2E: Add figure caption, update code chunk options

ggplot(data = ces, aes(x = white_percent, y = ces_3_0_score)) +
  geom_hex() +
  scale_fill_gradient(low = "thistle1", high = "slateblue4") +
  theme_minimal() +
  labs(x = "White Population (%)", 
       y = "Pollution Burden Score",
       title = "Cumulative Pollution Burden for White Population in California") +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill = guide_legend("Count")) # update legend title
```

**Figure 1.** CalEnviroScreen 3.0 aggregate score of pollution burden and population characteristics for the white demographic in California. Data: 2010 United States Census and Office of Environmental Health Hazard Assessment (OEHHA) at the California Environmental Protection Agency (CalEPA).
